CTL: Added value of differential correlation

### 2012-02-07 inst/doc/article.Rnw
CTL in inst/docs

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Danny Arends
% The qcl mapping routine
%\VignetteIndexEntry{Article}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article}

@@

\title{QCL: Added value of differential correlation}

@@

<<echo=FALSE,results=hide>>=
  require(qcl)
  set.seed(200)
  data(ath.result)
@

@@

\begin{sciabstract}
  In this paper we present Quanitative Correlation Locus (QCL) mapping, a novel approach 
  to detect genetic regulation of phenotypes in recombinant inbred line populations (RIL) 
  or SNP genotypes populations. QCL mapping allows association of phenotype correlation 
  differences to genetic markers, in almost the same way as QTL mapping associates genetic 
  markers with differences in expression.
  We use QCL mapping to reconstruct underlying genetic architecture, and provide insight 
  into regulation of (gene) expression. QCL information in combination with classical QTL 
  information leads to a more detailed insight into the genetic architecture underlying 
  complex traits. The method also provides a missing link in regulatory network reconstruction.
\end{sciabstract}

@@

  Recently different methods have been proposed that allow studying, in addition to the genes 
  differentially expressed between conditions, the genes that are differentially coexpressed 
  between conditions \cite{Tesson:2010,Kompass:2011}. Such differential co-expression 
  is taken as another level of information from genomic data. This level of information can for 
  example be used to expose transcriptional network rewiring \cite{Langfelder:2008}. Differential 
  co-expression has been observed in studies of different diseases, including between healthy and 
  cancer tissue studies. In parallel, correlations between phenotypes are studied for various 
  reasons e.g. discovering and detecting batch effects, marker assisted breeding and causal 
  inference. The importance of identifying differences in correlation has been increasingly 
  recognized and this led to the development of this algorithm to perform QCL analysis in 
  experimental crosses.

@@

  This inspired us to re-think parts of the basic assumptions in genetics:
  Imagine phenotype A shows a QTL, we suppose that near this marker something (e.g. DNA variation, 
  methylation, promotor sequences, etc) introduces expression variation in A. If we now assume there is a 
  phenotype B (downstream to A) variation of expression in phenotype A is causing 'variation' 
  for phenotype B. However B will in most cases not show variation because it's expression 
  depends (but is not dependant) on A, a phenomena generally refered to as phenotype buffering. 
  This is what QCL mapping exploits, the hypothesis is that although B doesn't show a QTL, 
  it will become uncorrelated when we partition phenotype B into the groups based on the genotype at 
  the QTL marker of phenotype A.

  QCL thus maps not differential co-expression but differential correlation to genetic 
  variation, i.e. to identify regions in the genome for which one genotype value leads to 
  correlated expression between a phenotype and all other phenotypes, exploiting the fact 
  that any QTL will be the result of either direct induced variation or indirect variation, 
  which will be visible as correlation differences among individuals carrying different genotypes. 
  We merge information coming from QCL and QTL to reveal underlying genetic architecture, 
  which would remain hidden when using only classical QTL information.

  We urge all QTL researchers to combine the new QCL method with existing QTL analysis to expose 
  transcriptional networks involved in the regulation of complex phenotypes. For this we 
  provide an R package with a reference implementation of the QCL algorithm.
\bigskip

@@

\emph{\sffamily QCL mapping}\\
  As mentioned in the introduction QCL mapping uses buffering to detect interaction partners. 
  When we split the population based on genotype at a QTL marker, only related phenotypes 
  will loose correlation with the gene under investigation. This allows us to not only 
  only identify its interaction partners, while not suffering from (p1) many phenotypes 
  co-localizing. Additionally it is more sensitive to detect variation 
  caused by regulatory phenotypes because variation of these regulators causes many 'downstream' 
  phenotypes to show differential correlation with the inducing phenotype, because QCL relies on 
  correlation changes as opposed to differences in phenotype mean QCL is unaffected by (p2).

\emph{\sffamily Choosing a start/seed probe (Optional)}\\
  Choose a phenotype to subject to QCL analysis, this can be your favourite gene, 
  protein, metabolite or classical phenotype.

\emph{\sffamily Single phenotype QCL mapping}\\
  Using genotype data we can split individuals at each marker in two groups (A or B).
  At each marker calculate the correlation between our phenotype and all other 
  phenotypes in groups A and B. Collect all the result vectors into two matrices A 
  and B, and calculate the QCL matrix by taking the squared difference between matrix A 
  and B.

\emph{\sffamily Significance testing}\\
  To compare to QTL mapping we need to convert the QCLscores to LOD scores. Significance 
  of QCL were assessed by using permutations, in each round the link between genotype and 
  phenotype is broken, by redistributing at random genotypes amongst individuals not 
  allowing for duplicates. The QCL on these new individuals allows for estimation of 
  the underlying null-distribution of QCL scores. We frequently observe QCL scores in 
  real data which is higher than any QCL score obtained during permutation. When this 
  happens we use a generalized Pareto distribution (GPD) to estimate the extreme tail 
  of the distribution \cite{Knijnenburg:2009}, to also get likelyhood estimates for the 
  extreme scores observed. See Fig. \ref{fig:permutation_scores} showing the score 
  distribution in permuted data of 3 traits in A. Thaliana, Fig XX also shows a tail 
  estimation using the GPD method.

\emph{\sffamily Creating single trait QCL profiles}\\
  After permutation single trait QCL profiles can be obtained by transforming the QCL scores 
  into LOD scores. For this we use a look-up table of scores obtained during permutation, we 
  assess the likelyhood and take the $-log10(P.value)$ of every QCL score see Fig. 
  \ref{fig:singletraitLOD}. This is done 1) to compare QTL profiles with summarized QCL profiles 
  (an example in Fig. \ref{fig:singletrait}), and 2) Show the contribution of each phenotype 
  to the total likelyhood.

\emph{\sffamily Visualizing the Phenotypes x Phenotypes QCL relations}\\
  In Fig. \ref{fig:singletraitLOD} we see the additional information obtained by QCL mapping summarized 
  graphically.  We can also calculate the full Phenotypes x Phenotypes matrix. This shows the summed 
  evidence of all paiwise interaction between multiple phenotypes.

\emph{\sffamily Visualizing the Phenotypes x Marker relations}\\
  When plotting multiple QCL profiles as a heatmap, QCLs are seen mapping to the 
  same genetic location. At these locations, regulation is taking place. We expect 
  to find one (or more) of the genes to show a cis-QTL, this is because a QCL can 
  only exist if there is a QTL inducing it.

@@

  We show two examples in which we reconstruct the genetic architecture using QCL mapping. 
  Both datasets are available inside the R package, as well as the code to analyse these 
  two datasets.

@@

  \subsection{Example 1. A. Thaliana}
  QCL mapping was performed on 24 characterized metabolities measured in Arabidopsis 
  Thaliana. A significance threshold for QCL mapping was determined for each trait 
  by using 1000 permutations, the outcomes can be seen in Fig \ref{fig:histPermAra}. 
  We used a very stringent QCL threshold at P < 0.001 to use as 'significant' QCL 
  score for further analysis.

@@

  \emph{\sffamily Single trait QCL}\\
  The 24 QCL profiles were compared to the QTL profiles previously determined. We observed 3 cases:\\
  1) Additional QCL found, where no QTL was present as seen in the QCL profile X3.Hydroxypropyl 
  (Fig. \ref{fig:singletraitLOD}), these are the cases in which the loss of variation in the genes 
  underneath the QCL is not causing variation differences in the observed phenotype, but still 
  we're able to detect the correlation differences.\\
  2) QCL co-localized with QTL, in this situation QCL gives a detailed view of which phenotypes 
  are underneath its peek, we interpret this as variation transfered from the phenotypes underneath 
  the QCL to the phenotype investigated.\\
  3) QTL with no QCL, we interpret this as the source of the variation. Analysis in this way of 
  all 24 Single trait QCL profiles combined with their QTL profiles allows us to reconstruct the 
  source of natural variation and follow it down the network.

  \emph{\sffamily QCL combined with QTL}\\
  Plots of the results of the QCL mapping in Arabidopsis can be found in 
  Figure XX Phenotype to Marker interaction, which is comparable to the QTL mapping 
  (Fig. XX). The additional information obtained by QCL mapping is shown in a heatmap 
  of the Phenotype to Phenotype interaction matrix (figure XX) and the same information 
  but now represented as an interaction network is seen in figure XX and in the QCL plot

  \emph{\sffamily Reconstructed network}\\
  Using the derived data we can begin to reconstruct the network. We'll start with a metabolite 
  which shows mappable variation changes, but no QCL colocatizing (Inducer).

  \subsection{Example 2. S. Cerevisiae}
  QCL mapping is computationaly expensive (We need 1000+ permutations for each phenotype), 
  and for the S. Cerevisiae we pre-selected a background of 300 probes which showed the 
  highest variation, and had been previously been identified as having QTLs. Significance 
  thresholds for QCL mapping were determined for each trait by using 1000 permutations 
  ({\emph Data not shown}).

  \emph{\sffamily Single trait QCL}\\

@@

  However results on complex traits have been less then satisfactory with only small 
  amounts of heritability explained by individuals loci. QCL mapping however 'borrows' 
  information from each phenotype in the analysis set, and enables extraction of 
  phenotype to marker to phenotype networks, the flow of variation in the network. 
  Furthermore QCLs seem to 'transfer' from phenotype to phenotypes in a stabile 
  fashion, which gives greater detail and confidence when reconstructing genetic 
  architecture. Also QCL allows us to better identify gene transcript introducing 
  variation, by their absense under QTL peeks which introduce variation, compared 
  to being present underneath QTL which are affected by upstream variation.

@@

%\begin{figure}[ht]
%  \centerline{\includegraphics[height=7cm]{qcl.pdf}}
%  \vspace{-15pt}
%  \caption{{\scriptsize QCL mapping: We start by looking at a single genetic marker and dividing our populaiton into two groups.
%  We start by choosing a phenotype (red), and use its expression to detect the red module (high correlation in gray genotype). 
%  We see that something underlying our current marker in the gray genotype regulates the expression of red (but not blue)}}
%  \vspace{-20pt}
%  \label{fig:qcl}
%\end{figure}

%\begin{figure}[ht]
%  \centerline{\includegraphics[height=6cm]{qcl_n.pdf}}
%  \vspace{-15pt}
%  \caption{{\scriptsize Reconstructed pathway location from gene expression data: Insight into the underlying (hidden) regulatory network 
%  can be obtained when we look at the amount of expression showing the same switch. Depending on the amount of blue expressions 
%  co-affected we can guestimate the location of the red expression inside the pathway}}
%  \vspace{-20pt}
%  \label{fig:qcl_n}
%\end{figure}

\begin{figure}[ht]
  \centering
<<permutation_scores, fig=TRUE, echo=FALSE, height=3.0>>=
op <- par(cex=0.5)
hist(ath.result,c(3,2,1))
op <- par(cex=1)
@ 
  \caption{{\scriptsize QCL scores obtained during permutations (A. Thaliana) additional information obtained by QCL mapping, 
  which can be used to reconstruct the underlying genetic architecture. }}
  \label{fig:histPermAra}
\end{figure}

\begin{figure}
  \centering
<<singletrait, fig=TRUE, echo=FALSE, height=3.0>>=
op <- par(cex=0.5)
plot(ath.result[[1]])
op <- par(cex=1)
@
  \caption{{\scriptsize QCL profile (A. Thaliana, X3.Hydroxypropyl) transformed to LOD scores compared to the QTL profile of the same phenotype, We see an 
  additional significant peek for this phenotype (Between Marker 50 to 60) }}
  \label{fig:singletrait}
\end{figure}

\begin{figure}
  \centering
<<singletraitLOD, fig=TRUE, echo=FALSE, height=6.0>>=
op <- par(cex=0.5)
QCLasLOD(ath.result[[1]], ath.result[[1]]$qtl)
op <- par(cex=1)
@ 
  \caption{{\scriptsize QCL profile (A. Thaliana, X3.Hydroxypropyl) transformed to LOD scores. Shown is the likelyhood contribution of each of 
  the 24 metabolites to the QCL profile.}}
  \label{fig:singletraitLOD}
\end{figure}

@@

\begin{figure}[ht]
  \centering
<<p2pmatrix, fig=TRUE, echo=FALSE, height=3.0>>=
par(mar=c(3,8,1.1,1.1))
plot(ath.result, against="phenotypes", do.grid=TRUE, grid.col="white")
@ 
  \caption{{\scriptsize Phenotypes x Phenotypes matrix (A. Thaliana) additional information obtained by QCL mapping, 
  which can be used to reconstruct the underlying genetic architecture}}
  \label{fig:p2pmatrix}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Danny Arends
% The ctl mapping routine
%\VignetteIndexEntry{Article}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{article}

@@

\title{CTL: Added value of differential correlation}

@@

<<echo=FALSE,results=hide>>=
  require(ctl)
  set.seed(200)
  data(ath.result)
@

@@

\begin{sciabstract}
  In this paper we present Quanitative Correlation Locus (CTL) mapping, a novel approach 
  to detect genetic regulation of phenotypes in recombinant inbred line populations (RIL) 
  or SNP genotypes populations. CTL mapping allows association of phenotype correlation 
  differences to genetic markers, in almost the same way as QTL mapping associates genetic 
  markers with differences in expression.
  We use CTL mapping to reconstruct underlying genetic architecture, and provide insight 
  into regulation of (gene) expression. CTL information in combination with classical QTL 
  information leads to a more detailed insight into the genetic architecture underlying 
  complex traits. The method also provides a missing link in regulatory network reconstruction.
\end{sciabstract}

@@

  Recently different methods have been proposed that allow studying, in addition to the genes 
  differentially expressed between conditions, the genes that are differentially coexpressed 
  between conditions \cite{Tesson:2010,Kompass:2011}. Such differential co-expression 
  is taken as another level of information from genomic data. This level of information can for 
  example be used to expose transcriptional network rewiring \cite{Langfelder:2008}. Differential 
  co-expression has been observed in studies of different diseases, including between healthy and 
  cancer tissue studies. In parallel, correlations between phenotypes are studied for various 
  reasons e.g. discovering and detecting batch effects, marker assisted breeding and causal 
  inference. The importance of identifying differences in correlation has been increasingly 
  recognized and this led to the development of this algorithm to perform CTL analysis in 
  experimental crosses.

@@

  This inspired us to re-think parts of the basic assumptions in genetics:
  Imagine phenotype A shows a QTL, we suppose that near this marker something (e.g. DNA variation, 
  methylation, promotor sequences, etc) introduces expression variation in A. If we now assume there is a 
  phenotype B (downstream to A) variation of expression in phenotype A is causing 'variation' 
  for phenotype B. However B will in most cases not show variation because it's expression 
  depends (but is not dependant) on A, a phenomena generally refered to as phenotype buffering. 
  This is what CTL mapping exploits, the hypothesis is that although B doesn't show a QTL, 
  it will become uncorrelated when we partition phenotype B into the groups based on the genotype at 
  the QTL marker of phenotype A.

  CTL thus maps not differential co-expression but differential correlation to genetic 
  variation, i.e. to identify regions in the genome for which one genotype value leads to 
  correlated expression between a phenotype and all other phenotypes, exploiting the fact 
  that any QTL will be the result of either direct induced variation or indirect variation, 
  which will be visible as correlation differences among individuals carrying different genotypes. 
  We merge information coming from CTL and QTL to reveal underlying genetic architecture, 
  which would remain hidden when using only classical QTL information.

  We urge all QTL researchers to combine the new CTL method with existing QTL analysis to expose 
  transcriptional networks involved in the regulation of complex phenotypes. For this we 
  provide an R package with a reference implementation of the CTL algorithm.
\bigskip

@@

\emph{\sffamily CTL mapping}\\
  As mentioned in the introduction CTL mapping uses buffering to detect interaction partners. 
  When we split the population based on genotype at a QTL marker, only related phenotypes 
  will loose correlation with the gene under investigation. This allows us to not only 
  only identify its interaction partners, while not suffering from (p1) many phenotypes 
  co-localizing. Additionally it is more sensitive to detect variation 
  caused by regulatory phenotypes because variation of these regulators causes many 'downstream' 
  phenotypes to show differential correlation with the inducing phenotype, because CTL relies on 
  correlation changes as opposed to differences in phenotype mean CTL is unaffected by (p2).

\emph{\sffamily Choosing a start/seed probe (Optional)}\\
  Choose a phenotype to subject to CTL analysis, this can be your favourite gene, 
  protein, metabolite or classical phenotype.

\emph{\sffamily Single phenotype CTL mapping}\\
  Using genotype data we can split individuals at each marker in two groups (A or B).
  At each marker calculate the correlation between our phenotype and all other 
  phenotypes in groups A and B. Collect all the result vectors into two matrices A 
  and B, and calculate the CTL matrix by taking the squared difference between matrix A 
  and B.

\emph{\sffamily Significance testing}\\
  To compare to QTL mapping we need to convert the CTLscores to LOD scores. Significance 
  of CTL were assessed by using permutations, in each round the link between genotype and 
  phenotype is broken, by redistributing at random genotypes amongst individuals not 
  allowing for duplicates. The CTL on these new individuals allows for estimation of 
  the underlying null-distribution of CTL scores. We frequently observe CTL scores in 
  real data which is higher than any CTL score obtained during permutation. When this 
  happens we use a generalized Pareto distribution (GPD) to estimate the extreme tail 
  of the distribution \cite{Knijnenburg:2009}, to also get likelyhood estimates for the 
  extreme scores observed. See Fig. \ref{fig:permutation_scores} showing the score 
  distribution in permuted data of 3 traits in A. Thaliana, Fig XX also shows a tail 
  estimation using the GPD method.

\emph{\sffamily Creating single trait CTL profiles}\\
  After permutation single trait CTL profiles can be obtained by transforming the CTL scores 
  into LOD scores. For this we use a look-up table of scores obtained during permutation, we 
  assess the likelyhood and take the $-log10(P.value)$ of every CTL score see Fig. 
  \ref{fig:singletraitLOD}. This is done 1) to compare QTL profiles with summarized CTL profiles 
  (an example in Fig. \ref{fig:singletrait}), and 2) Show the contribution of each phenotype 
  to the total likelyhood.

\emph{\sffamily Visualizing the Phenotypes x Phenotypes CTL relations}\\
  In Fig. \ref{fig:singletraitLOD} we see the additional information obtained by CTL mapping summarized 
  graphically.  We can also calculate the full Phenotypes x Phenotypes matrix. This shows the summed 
  evidence of all paiwise interaction between multiple phenotypes.

\emph{\sffamily Visualizing the Phenotypes x Marker relations}\\
  When plotting multiple CTL profiles as a heatmap, CTLs are seen mapping to the 
  same genetic location. At these locations, regulation is taking place. We expect 
  to find one (or more) of the genes to show a cis-QTL, this is because a CTL can 
  only exist if there is a QTL inducing it.

@@

  We show two examples in which we reconstruct the genetic architecture using CTL mapping. 
  Both datasets are available inside the R package, as well as the code to analyse these 
  two datasets.

@@

  \subsection{Example 1. A. Thaliana}
  CTL mapping was performed on 24 characterized metabolities measured in Arabidopsis 
  Thaliana. A significance threshold for CTL mapping was determined for each trait 
  by using 1000 permutations, the outcomes can be seen in Fig \ref{fig:histPermAra}. 
  We used a very stringent CTL threshold at P < 0.001 to use as 'significant' CTL 
  score for further analysis.

@@

  \emph{\sffamily Single trait CTL}\\
  The 24 CTL profiles were compared to the QTL profiles previously determined. We observed 3 cases:\\
  1) Additional CTL found, where no QTL was present as seen in the CTL profile X3.Hydroxypropyl 
  (Fig. \ref{fig:singletraitLOD}), these are the cases in which the loss of variation in the genes 
  underneath the CTL is not causing variation differences in the observed phenotype, but still 
  we're able to detect the correlation differences.\\
  2) CTL co-localized with QTL, in this situation CTL gives a detailed view of which phenotypes 
  are underneath its peek, we interpret this as variation transfered from the phenotypes underneath 
  the CTL to the phenotype investigated.\\
  3) QTL with no CTL, we interpret this as the source of the variation. Analysis in this way of 
  all 24 Single trait CTL profiles combined with their QTL profiles allows us to reconstruct the 
  source of natural variation and follow it down the network.

  \emph{\sffamily CTL combined with QTL}\\
  Plots of the results of the CTL mapping in Arabidopsis can be found in 
  Figure XX Phenotype to Marker interaction, which is comparable to the QTL mapping 
  (Fig. XX). The additional information obtained by CTL mapping is shown in a heatmap 
  of the Phenotype to Phenotype interaction matrix (figure XX) and the same information 
  but now represented as an interaction network is seen in figure XX and in the CTL plot

  \emph{\sffamily Reconstructed network}\\
  Using the derived data we can begin to reconstruct the network. We'll start with a metabolite 
  which shows mappable variation changes, but no CTL colocatizing (Inducer).

  \subsection{Example 2. S. Cerevisiae}
  CTL mapping is computationaly expensive (We need 1000+ permutations for each phenotype), 
  and for the S. Cerevisiae we pre-selected a background of 300 probes which showed the 
  highest variation, and had been previously been identified as having QTLs. Significance 
  thresholds for CTL mapping were determined for each trait by using 1000 permutations 
  ({\emph Data not shown}).

  \emph{\sffamily Single trait CTL}\\

@@

  However results on complex traits have been less then satisfactory with only small 
  amounts of heritability explained by individuals loci. CTL mapping however 'borrows' 
  information from each phenotype in the analysis set, and enables extraction of 
  phenotype to marker to phenotype networks, the flow of variation in the network. 
  Furthermore CTLs seem to 'transfer' from phenotype to phenotypes in a stabile 
  fashion, which gives greater detail and confidence when reconstructing genetic 
  architecture. Also CTL allows us to better identify gene transcript introducing 
  variation, by their absense under QTL peeks which introduce variation, compared 
  to being present underneath QTL which are affected by upstream variation.

@@

%\begin{figure}[ht]
%  \centerline{\includegraphics[height=7cm]{CTL.pdf}}
%  \vspace{-15pt}
%  \caption{{\scriptsize CTL mapping: We start by looking at a single genetic marker and dividing our populaiton into two groups.
%  We start by choosing a phenotype (red), and use its expression to detect the red module (high correlation in gray genotype). 
%  We see that something underlying our current marker in the gray genotype regulates the expression of red (but not blue)}}
%  \vspace{-20pt}
%  \label{fig:CTL}
%\end{figure}

%\begin{figure}[ht]
%  \centerline{\includegraphics[height=6cm]{CTL_n.pdf}}
%  \vspace{-15pt}
%  \caption{{\scriptsize Reconstructed pathway location from gene expression data: Insight into the underlying (hidden) regulatory network 
%  can be obtained when we look at the amount of expression showing the same switch. Depending on the amount of blue expressions 
%  co-affected we can guestimate the location of the red expression inside the pathway}}
%  \vspace{-20pt}
%  \label{fig:CTL_n}
%\end{figure}

\begin{figure}[ht]
  \centering
<<permutation_scores, fig=TRUE, echo=FALSE, height=3.0>>=
op <- par(cex=0.5)
hist(ath.result,c(3,2,1))
op <- par(cex=1)
@ 
  \caption{{\scriptsize CTL scores obtained during permutations (A. Thaliana) additional information obtained by CTL mapping, 
  which can be used to reconstruct the underlying genetic architecture. }}
  \label{fig:histPermAra}
\end{figure}

\begin{figure}
  \centering
<<singletrait, fig=TRUE, echo=FALSE, height=3.0>>=
op <- par(cex=0.5)
plot(ath.result[[1]])
op <- par(cex=1)
@
  \caption{{\scriptsize CTL profile (A. Thaliana, X3.Hydroxypropyl) transformed to LOD scores compared to the QTL profile of the same phenotype, We see an 
  additional significant peek for this phenotype (Between Marker 50 to 60) }}
  \label{fig:singletrait}
\end{figure}

\begin{figure}
  \centering
<<singletraitLOD, fig=TRUE, echo=FALSE, height=6.0>>=
op <- par(cex=0.5)
CTLasLOD(ath.result[[1]], ath.result[[1]]$qtl)
op <- par(cex=1)
@ 
  \caption{{\scriptsize CTL profile (A. Thaliana, X3.Hydroxypropyl) transformed to LOD scores. Shown is the likelyhood contribution of each of 
  the 24 metabolites to the CTL profile.}}
  \label{fig:singletraitLOD}
\end{figure}

@@

\begin{figure}[ht]
  \centering
<<p2pmatrix, fig=TRUE, echo=FALSE, height=3.0>>=
par(mar=c(3,8,1.1,1.1))
plot(ath.result, against="phenotypes", do.grid=TRUE, grid.col="white")
@ 
  \caption{{\scriptsize Phenotypes x Phenotypes matrix (A. Thaliana) additional information obtained by CTL mapping, 
  which can be used to reconstruct the underlying genetic architecture}}
  \label{fig:p2pmatrix}
\end{figure}
